Framing active school travel in Ontario, or how spinach is good for you

### 2023-05-26 AST-Framing-Ontario/AST-Framing-Ontario.tex
Resubmission to TR-A after revisions by ED, DC, DR, EOW, and AP

\usepackage{lineno} % add

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={How is active school travel framed in Ontario, Canada?},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{How is active school travel framed in Ontario, Canada?}
    \author[McMaster University]{Elise Desjardins%
  %
  }
   \ead{desjae@mcmaster.ca} 
    \author[McMaster University]{Jason Lam%
  %
  }
   \ead{lamj54@mcmaster.ca} 
    \author[University of Alberta]{Darcy Reynard%
  %
  }
   \ead{reynard@ualberta.ca} 
    \author[University of Alberta]{Damian Collins%
  %
  }
   \ead{damian.collins@ualberta.ca} 
    \author[Polytechnique Montreal]{E. Owen D. Waygood%
  %
  }
   \ead{owen.waygood@polymtl.ca} 
    \author[McMaster University]{Antonio Paez%
  \corref{cor1}%
  \fnref{1}}
   \ead{paezha@mcmaster.ca} 
      \affiliation[McMaster University]{School of Earth, Environment and
Society, 1280 Main St.~W, Hamilton, Ontario, Canada L8S 4K1}
    \affiliation[University of Alberta]{Department of Earth and
Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta,
Canada T6G 2E3}
    \affiliation[Polytechnique Montreal]{Department of Civil, Geological
and Mining Engineering, 2500, chemin de Polytechnique Montreal,
Montreal, Quebec, Canada H3T 1J4}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding author.}

  \begin{abstract}
  Active school travel (AST) has been increasingly encouraged by various
  stakeholders in Ontario, Canada through efforts such as school travel
  planning. Education strategies like workshops or resources that
  promote AST are commonly implemented. The framing of AST through such
  strategies may influence how walking and bicycling to school are
  perceived by parents. It may also draw attention to AST as an issue
  affecting children's health which could motivate behaviour change. We
  used natural language processing, including topic modelling, to
  examine how AST is framed in publicly available documents from Ontario
  stakeholders involved in school travel planning. We then compared the
  findings from these documents to a selection of studies on AST and
  explored similarities between the two. We found that AST is framed in
  two ways: i) as a health and environmental issue; and ii) as an
  accessible and feasible transport option for children and parents. The
  frames encourage children and parents to adopt AST given its health
  and environmental benefits by providing resources to support behaviour
  change. The benefits of AST and strategies to support AST that are
  communicated by stakeholders are consistent with the evidence from
  academic research. While these frames present AST in a positive light,
  they may not encourage parents to view current household travel
  behaviours as unhealthy for their children or their community.
  Stakeholders promoting AST in Ontario should further problematize the
  decline of AST and challenge the norm of driving children to school.
  \end{abstract}
    \begin{keyword}
    active school travel \sep school travel planning \sep natural
language processing \sep 
    framing analysis
  \end{keyword}

@@

Walking and bicycling to school, commonly known as active school travel
(AST), have declined in North America for decades
\citep{rothmanDeclineActiveSchool2018}, but the appetite for these modes
is beginning to rise in many Canadian communities. For instance,
research in Toronto, Ontario found that 40\% of children who were driven
to school would like to travel by bicycle instead
\citep{laroucheRatherBikeSchool2016}. Another study in London, Ontario,
reported similar findings with respect to children's preference for
active travel \citep{larsenRouteBasedAnalysisCapture2012}. From a public
health perspective, increasing active travel to school represents a
major opportunity to improve children's health and wellbeing.

To promote this trend, in 2017 the Government of Ontario began to
support communities develop AST initiatives through a program called
\emph{Ontario Active School Travel} (OAST). By 2021, the program, run by
Green Communities Canada had awarded over CAD 2 million and provided
resources to over 25 projects across Ontario
\citep{GreenCommunities2016}. A popular intervention in many communities
is school travel planning (STP) \citep[see][]{GreenCommunitiesProjects}.
In this context, STPs are ``school-specific'' interventions led by a
facilitator who works with a committee of stakeholders from diverse
sectors including education, planning, transportation, and public health
to develop action plans
\citep{buliungSchoolTravelPlanning2011, mammenSchoolTravelPlanning2014}.
The five-step process involves identifying barriers to AST and
implementing solutions or activities to make AST safer and more
convenient: 1) setup of the program; 2) data collection and problem
identification; 3) action planning; 4) implementation; 5) monitoring and
evaluation
\citep{buliungSchoolTravelPlanning2011, langUnderstandingModalChoice2011}.

Framing AST by STP stakeholders to a target audience (e.g., parents or
the general public) can raise awareness of the relevant issues and
influence how active travel to school is perceived. Gamson and
Modigliani \citeyearpar{gamsonChangingCulture1987} define a frame as a
``central organizing idea or story line that provides meaning'' to a
particular phenomenon. A frame can enable individuals ``to locate,
perceive, identify, and label'' information pertaining to various
dimensions of an issue \citep{goffmanFrameAnalysisEssay1974}. Therefore,
stakeholders involved in STP efforts can play a role in shaping public
perception about AST to attract greater attention and recognition as a
``problem'' that needs to be addressed through behaviour change or new
policies.

After significant investment of human and financial resources to boost
rates of AST in Ontario over the past few years, in this paper we ask
this central question: How is AST framed to the public in Ontario? The
following questions are subsidiary and help us address this main
question: What benefits of AST are communicated to the public? What
solutions are proposed to increase AST?

We approach this question by means of natural language processing, and
examine how three STP stakeholder groups in Ontario (municipalities,
school boards, and transportation consortia) frame AST. We assemble a
corpus of texts from public sources that present information for the
general public or parents who are interested in AST. We examine word
frequency, bigrams, and concordances in these selected documents, and
also identify key topics presented by each stakeholder group. We then
compare the findings from these documents to a selection of studies on
AST and explore the extent to which there is concordance between the
scholarly literature on AST and materials shared with the public.

@@

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM) \citep{mitraIndependentMobilityMode2013} or
systems model \citep{badlandDevelopmentSystemsModel2016} whereby
children's travel behaviour is understood within the context of
household, social, neighbourhood, and policy environments. The SEM helps
to identify multiple determinants that need to be addressed by
school-based interventions to increase AST, with the consensus being
that interventions should target multiple levels
\citep{mitraIndependentMobilityMode2013}. Rather than describing an
extensive list of factors that influence AST and mode choice in Canada
\citep[e.g.,][]{mammenUnderstandingDriveEscort2012, mitraIndependentMobilityMode2013, rothmanDeclineActiveSchool2018, wilsonUnderstandingChildParent2018},
we discuss a few potentially modifiable determinants that may be
targeted for change through STP.

Children's mode choice to school is strongly influenced by their
parents' travel behaviours and the complexity of their household's
travel needs \citep{buliungLivingJourneySchool2021}, as well as daily
parental support \citep{mahDoesParentalSupport2017a}. This indicates
that shifting parental perceptions and habits is important. Convenience
and inclement weather have been cited by parents as barriers to AST
\citep{buliungSchoolTravelPlanning2011}. Parental perceptions of the
built or school environment
\citep{demeesterParentalPerceivedNeighborhood2014, panterAttitudesSocialSupport2010}
and their children's skills
\citep{mammenUnderstandingDriveEscort2012, faulknerWhatQuickestEasiest2010}
also influence mode choice to school.

Distance between home and school is most negatively associated with AST
\citep{ikedaAssociationsChildrenActive2018, mammenUnderstandingDriveEscort2012, pontEnvironmentalCorrelatesChildren2009, rothmanDeclineActiveSchool2018}
with less AST reported among children who have to travel farther to
school. Many studies have also found that the quality of the built
environment along the route to school and around the school site
\citep{ikedaAssociationsChildrenActive2018, rothmanActiveSchoolTransportation2021}
and provision of active travel infrastructure
\citep{chenPromotingActiveStudent2018, pontEnvironmentalCorrelatesChildren2009}
facilitate AST. Canadian youth report that they feel most safe bicycling
on streets in their neighbourhood or that have low volumes of traffic
\citep{TACbikeinfra2020}. Finally, concerns about traffic and strangers
have been reported by parents who drive their children to school
\citep{mammenUnderstandingDriveEscort2012}, which highlights that the
volume and speed of cars can be a concern or deterrent for AST.

\hypertarget{school-travel-planning-in-canada-and-relevance-of-framing-analysis}{%
\section{School travel planning in Canada and relevance of framing
analysis}\label{school-travel-planning-in-canada-and-relevance-of-framing-analysis}}

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators establish
multi-sector committees which intervene at the participating school
through a range of activities consisting of \emph{{[}E{]}ducation}
strategies, \emph{{[}E{]}ncouragement} through in-person events or
programs, \emph{{[}E{]}ngineering} improvements to or around the school
site, and \emph{{[}E{]}nforcement} of traffic speeds around schools
\citep[or
4Es,][]{langUnderstandingModalChoice2011, mammenActiveSchoolTravel2014}.
Following the first large-scale evaluation of STP as an intervention in
Canada, AST rates increased and 14\% of parents reported a perceived
reduction in traffic outside the school
\citep{buliungSchoolTravelPlanning2011}. Assessments of the efficacy of
STP
\citep{buttazzoniPromotingActiveSchool2019, mammenActiveSchoolTravel2014}
indicate its potential to encourage behaviour change and adoption of
AST.

By providing a ``central organizing idea or story line''
\citep{gamsonChangingCulture1987} framing of AST seems particularly
important to shift parental attitudes and perceptions, given their
reported influence on children's travel mode to school. Scholars in the
field of political communications have proposed that communicators, such
as the media or institutions, construct the narrative of a frame for
policy positions or public issues in order to activate or restrict a
particular response in the intended audience
\citep{panFramingAnalysisApproach1993}. Organized groups of stakeholders
can employ similar methods to attract attention to particular issues.
Framing can be used to position existing solutions as suitable to
address particular issues \citep{mahFramecriticalPolicyAnalysis2014},
which may prevent the public from being aware of other policy approaches
that challenge the status quo. The way policy issues are framed is
ultimately important to understand because it plays a role in either
altering or preserving the existing social perceptions. Finally, the
ideas that are communicated by a particular frame can be either positive
or negative \citep{waygoodCO2ValenceFraming2018}. There is evidence that
negative framing may be more effective at motivating individuals, for
example, to change transport modes to reduce CO2 emissions
\citep{waygoodCO2ValenceFraming2018}.

Framing of AST could also affect broader support in the community.
Municipal representatives are perceived to be instrumental but the
involvement of other stakeholder groups (e.g., busing consortia
representatives and local residents) can be lacking
\citep{buttazzoniSupportingActiveSchool2018}. In Ontario, it is
reasonable to say that AST has become a policy issue on the education
and public health agendas, supported, albeit in a modest way, by
financial contributions from the provincial government. Still, the
support from a range of municipal representatives
\citep[see][]{buttazzoniSupportingActiveSchool2018, mammenPuttingSchoolTravel2015}
demonstrates that the AST issue is on the political agenda.

Since STP activities heavily focus on education or encouragement
\citep{buliungSchoolTravelPlanning2011, mammenActiveSchoolTravel2014, buttazzoniSupportingActiveSchool2018},
it is important that parents and communities are receptive to the goals
of STP \citep{buttazzoniSupportingActiveSchool2018}. Parents have been
found to express different understandings, language, and perceptions
than planners of how the built environment can influence school travel
\citep{buliungLivingJourneySchool2021}. This is also true when it comes
to other factors like convenience of different modes to school
\citep{langUnderstandingModalChoice2011}. STP stakeholders must pay
special attention to parents' understanding of the decline of AST as a
problem, which may affect their receptivity to proposed solutions.

The success of STP interventions would likely depend on parental
judgments of factors that are related to school travel, as well as
support from key policy makers. For example, parents have been found to
view mixed land use as conducive for driving, despite transport planners
viewing neighbourhoods with mixed uses as key for encouraging more
active travel \citep{buliungLivingJourneySchool2021}. Therefore,
stakeholders involved in STP must make conscious choices about the
proposed solutions and potential benefits of AST that are communicated
to parents and the general public through publicly available content to
convey its importance as a policy issue and to facilitate adoption of
AST.

@@

We conducted a search on Web of Science for scholarly papers on the
topic of school travel. We limited our search to the fields of public
health, transportation, planning, urban studies, and geography. This
initial corpus was curated by the authors to ensure that all documents
were relevant. After this step, 227 journal articles were readied for
analysis.

@@

We use topic modelling to analyze the documents retrieved. Topic
modelling is a natural language processing technique used to analyze
text to identify the language and concepts being communicated. This
method is practical for researchers working with large amounts of text
because it replaces the manual coding of topics that would normally take
place to analyze or summarize textual data
\citep{jacobiQuantitativeAnalysisLarge2016}. We primarily use the
following software: \texttt{tidytext} \citep{R-tidytext},
\texttt{topicmodels} \citep{R-topicmodels}, \texttt{word2vec}
\citep{R-word2vec}, and \texttt{wordcloud} \citep{R-wordcloud}.

@@

This paper is an example of open and reproducible research that uses
only open software. Following best practices in spatial data science
\citep{brunsdon2020opening}, the code and data needed to reproduce our
research or conduct a similar analysis for other regions are available
for download upon request.

@@

We analyzed word and document frequency for each corpus of text. Table
\ref{tab:word-table} shows the most frequent terms found in the
municipal, transportation consortia, school board, and academic
documents. Policy documents and academic papers reference \emph{active},
\emph{travel}, \emph{walking}, \emph{biking} or \emph{cycling}, and
\emph{students} more than other terms. Each corpus also has
\emph{safety} and \emph{traffic} as common words which suggests that
these are common concerns for parents. The word \emph{physical} is
present in each corpus, but it's not clear what this refers to (e.g,
\emph{physical activity}, or the \emph{physical environment}).
Furthermore, documents from STP stakeholder groups discuss
\emph{resources}, \emph{information}, and \emph{services} about school
travel. Unlike the academic papers, policy+practice documents include
the words \emph{route} or \emph{routes} as frequent terms. This could
indicate the role of STP stakeholder groups in identifying safe routes
to school to share with parents or families. In the section below, the
context in which these terms are used is explored further.

The academic corpus differs from the policy documents in that
\emph{parents} and \emph{distance} are the second and third most common
terms. In addition, \emph{time}, \emph{factors}, \emph{environment}, and
\emph{age} are also identified in the academic papers. The prevalence of
these terms is consistent with an academic focus on exploring the
variables that influence mode choice. These words are absent from the
list of common words in policy+practice documents. Table
\ref{tab:word-table} indicates that the academic corpus discusses a
broader range of determinants of AST than the policy documents. The
number of references for each term in the academic papers is also
substantially higher due to the inclusion of more documents.

@@

Examination of document frequency reveals terms that are not present in
all policy documents. This suggests that although documents pertain to
the subject of school travel, not all stakeholders across Ontario
disseminate information about AST. We manually searched the
policy+practice corpus and found that 48\% of documents mention AST and
16\% mention STP. In contrast, inclement weather and its impacts on
busing is a common topic addressed in school board and transportation
consortia documents.

@@

Bigrams refer to a pair of consecutive words. We combined all
municipality, school board, and transportation consortia documents into
the policy+practice. Figure \ref{fig:policy-visual} shows all of the
bigrams that occur more than 10 times in the policy+practice corpus.
This figure highlights the main ideas that are presented to the public
in each of the policy corpora. The directional arrows indicate the
arrangement of the words (e.g., active travel and not travel active) and
the colour gradient of the arrows corresponds to the most frequently
mentioned pairs (e.g., bigrams with darker arrows are found in the
corpus more often).

>>>>

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Framing active school travel in Ontario, or how spinach is good for you},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{Framing active school travel in Ontario, or how spinach is good
for you}
    \author[McMaster University]{Elise Desjardins%
  %
  }
   \ead{desjae@mcmaster.ca} 
    \author[McMaster University]{Jason Lam%
  %
  }
   \ead{lamj54@mcmaster.ca} 
    \author[University of Alberta]{Darcy Reynard%
  %
  }
   \ead{reynard@ualberta.ca} 
    \author[University of Alberta]{Damian Collins%
  %
  }
   \ead{damian.collins@ualberta.ca} 
    \author[Polytechnique Montreal]{E. Owen D. Waygood%
  %
  }
   \ead{owen.waygood@polymtl.ca} 
    \author[McMaster University]{Antonio Paez%
  \corref{cor1}%
  \fnref{1}}
   \ead{paezha@mcmaster.ca} 
      \affiliation[McMaster University]{School of Earth, Environment and
Society, 1280 Main St.~W, Hamilton, Ontario, Canada L8S 4K1}
    \affiliation[University of Alberta]{Department of Earth and
Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta,
Canada T6G 2E3}
    \affiliation[Polytechnique Montreal]{Department of Civil, Geological
and Mining Engineering, 2500, chemin de Polytechnique Montreal,
Montreal, Quebec, Canada H3T 1J4}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding author.}

  \begin{abstract}
  Active school travel (AST) is promoted in many jurisdictions,
  including Ontario, Canada, where there is provincial support for
  school travel planning (STP) efforts. Two pillars of school travel
  plans in the province are education and encouragement. For these
  pillars to stand, relevant stakeholders must be cognizant of and
  understand the issues and stakes. For this reason, we deem it
  important to understand how AST is communicated. In this research we
  adopt framing analysis to investigate the ways in which various
  organizations communicate to the public around AST. A frame is a
  central organizing idea or story line that provides meaning to a
  particular phenomenon, and thus helps to set the parameters for
  conversations about policy. In the case of AST, framing can influence
  what policy alternatives are perceived as available by children,
  parents, and their wider communities. The research is supported by
  natural language processing techniques applied to publicly available
  documents from Ontario stakeholders involved in school travel
  planning. We then compare the findings from these documents to a
  selection of academic studies on AST. We conclude that framing of AST
  in Ontario is mostly empirical-scientific in style, and largely in
  agreement with academic research on AST. However, it is essentially
  conservative and does challenge to the status quo of motorized travel.
  Furthermore, the frames tend to download responsibility for change to
  households, which limits the scope of policy alternatives by keeping
  collective action out of the frame.
  \end{abstract}
    \begin{keyword}
    active school travel \sep school travel planning \sep natural
language processing \sep 
    framing analysis
  \end{keyword}

@@

Walking and bicycling to school, commonly known as active school travel
(AST), have declined in North America for decades
\citep{rothmanDeclineActiveSchool2018}, but the appetite for these modes
is beginning to rise in many Canadian communities. For instance,
research in Toronto, Ontario found that 40\% of children who were driven
to school would like to travel by bicycle instead
\citep{laroucheRatherBikeSchool2016}. Another study in London, Ontario,
reported similar findings with respect to children's preference for
active travel \citep{larsenRouteBasedAnalysisCapture2012}. From a public
policy perspective, increasing active travel to school presents major
opportunities to improve children's health and wellbeing
\citep{faulknerActiveSchoolTransport2009}, to mitigate traffic and
improve perceptions of safety
\citep{rothmanAssociationsParentsPerception2015}, as well as to reduce
emissions and other forms of pollution
\citep{bearman2014modelling, goodman2019scenarios}.

To promote this trend, in 2017 the Government of Ontario began to
support communities develop AST initiatives through a program called
\emph{Ontario Active School Travel} (OAST). By 2021, the program, run by
\emph{Green Communities Canada} had awarded over CAD 2 million and
provided resources to over 25 projects across Ontario
\citep{GreenCommunities2016}. A popular intervention in many communities
is school travel planning (STP) \citep[see][]{GreenCommunitiesProjects}.
In this context, STPs are ``school-specific'' interventions led by a
facilitator who works with a committee of stakeholders from diverse
sectors including education, planning, transportation, and public health
to develop action plans
\citep{buliungSchoolTravelPlanning2011, mammenSchoolTravelPlanning2014}.
The five-step process involves identifying barriers to AST and
implementing solutions or activities to make AST safer and more
convenient: 1) setup of the program; 2) data collection and problem
identification; 3) action planning; 4) implementation; 5) monitoring and
evaluation \citep{hinckson2018school}.

How AST is presented to a target audience (e.g., parents or the general
public) is essential to define the parameters of what is possible
policy-wise. Drawing from concepts in communication studies, the
objective of this paper is to investigate how AST is framed in Ontario.
Gamson and Modigliani \citeyearpar{gamsonChangingCulture1987} define a
frame as a ``central organizing idea or story line that provides
meaning'' to a particular phenomenon. A frame can enable individuals
``to locate, perceive, identify, and label'' information pertaining to
various dimensions of an issue \citep{goffmanFrameAnalysisEssay1974}.
From this perspective, stakeholders involved in STP efforts play an
important role in shaping public perceptions about AST by generating
frames that create meaning, and thus influence the degree of attention
and recognition of a ``problem'' that needs to be addressed through
behavior change or new policies.

Our research is supported by natural language processing techniques,
applied to the analysis of documents produced by three STP stakeholder
groups in Ontario (municipalities, school boards, and transportation
consortia). We assemble a corpus of texts from public sources that
present information for the general public or parents who are interested
in AST and examine word frequency, bigrams, and concordances in these
selected documents. Furthermore, we use topic analysis to identify
themes presented by each stakeholder group. We then compare the findings
from these documents to a selection of studies on AST and explore the
extent to which there is concordance between the scholarly literature on
AST and materials shared with the public. The results of this analysis,
along with manual checks of the corpus by the authors, lead us to
conclude that AST is framed as a matter of individual choices instead of
collective action. The message is reasonable but inherently conservative
and it fails to communicate a sense of urgency about health and
environmental issues, which may help to explain the modest success of
related policies
\citep{buliungSchoolTravelPlanning2011, mammenSchoolTravelPlanning2014, mammenActiveSchoolTravel2014}

@@

Rather than describing an extensive list of factors that influence AST
and mode choice for travel to school in Canada
\citep[e.g.,][]{mammenUnderstandingDriveEscort2012, mitraIndependentMobilityMode2013, rothmanDeclineActiveSchool2018, wilsonUnderstandingChildParent2018},
we focus on two key aspects that may be targeted from a policy
perspective: built environmental factors and perceptions and behaviors
of parents.

The built environment is a key factor for physical planning, and in this
respect distance between home and school is the factor most negatively
associated with AST
\citep{ikedaAssociationsChildrenActive2018, mammenUnderstandingDriveEscort2012, pontEnvironmentalCorrelatesChildren2009, rothmanDeclineActiveSchool2018}
with less AST reported among children who have to travel farther to
school. Many studies have also found that the quality of the built
environment along the route to school and around the school site
\citep{ikedaAssociationsChildrenActive2018, rothmanActiveSchoolTransportation2021}
and provision of active travel infrastructure
\citep{chenPromotingActiveStudent2018, pontEnvironmentalCorrelatesChildren2009}
facilitate AST. Canadian youth report that they feel most safe bicycling
on streets in their neighbourhood or that have low volumes of traffic
\citep{TACbikeinfra2020}. Finally, concerns about traffic and strangers
have been reported by parents who drive their children to school
\citep{mammenUnderstandingDriveEscort2012}: ironically, concerns about
the dangers of traffic become a vicious cycle
\citep{rothmanSchoolEnvironmentStudent2017}.

Secondly, research has found that parents are key decision-makers with
respect to children's mode of travel to school
\citep{rothmanAssociationsParentsPerception2015}. Particularly, after
distance \citep{curtis2015built}, parental perceptions of the built or
school environment, as well as of their children's skills, have been
found to be key in influencing mode choice to school
\citep{demeesterParentalPerceivedNeighborhood2014, panterAttitudesSocialSupport2010, mandic2020differences, mammenUnderstandingDriveEscort2012, faulknerWhatQuickestEasiest2010}.
For example, \citet{ramanathanHappinessMotionEmotions2014a} found that
parents and children who perceived active travel as beneficial for
health and wellbeing benefits were more likely to use active modes to
get to school. Furthermore, children's mode choice to school is strongly
influenced by their parents' travel behaviors, the complexity of their
household's travel needs \citep{buliungLivingJourneySchool2021}, as well
as daily parental support \citep{mahDoesParentalSupport2017a}. This
suggests that shifting parental perceptions and habits is important.

\hypertarget{stp-frames}{%
\section{School travel planning in Canada and relevance of
frames}\label{stp-frames}}

AST has become a policy issue on the education and public health agendas
in Ontario, and is supported, albeit in a modest way, by financial
contributions from the provincial government. Still, endorsement from a
range of municipal representatives
\citep[see][]{buttazzoniSupportingActiveSchool2018, mammenPuttingSchoolTravel2015}
demonstrates that the AST issue is on the political agenda. One of the
main vehicles to promote AST is school travel planning (STP), which has
been implemented in Canada since at least the late 2000s. Within the STP
process, facilitators establish multi-sector committees which intervene
at the participating school through four pillars consisting of
\emph{{[}E{]}ducation} strategies, \emph{{[}E{]}ncouragement} through
in-person events or programs, \emph{{[}E{]}ngineering} improvements to
or around the school site, and \emph{{[}E{]}nforcement} of traffic
speeds around schools \citep[or
4Es,][]{langUnderstandingModalChoice2011, mammenActiveSchoolTravel2014}.
Of the 4Es, {[}E{]}ngineering targets the first set of factors discussed
in the preceding section (the built environment), whereas
{[}E{]}ducation and {[}E{]}ncouragement target perceptions.

The relevance of policy frames with respect to education and
encouragement is highlighted by \citet{beland2014developing}, who
remarks ``framing processes as public relations are central to policy
issues related to sustainable transportation''. Seeing how influential
parental perceptions are to AST, it is important to understand
AST-related communications, since the construction of frames for policy
positions or public issues can activate or restrict a particular
response in the intended audience
\citep{panFramingAnalysisApproach1993}. Besides parents and children,
framing can also affect broader policy support for AST in the community.
Municipal representatives are perceived to be instrumental but the
involvement of other stakeholder groups (e.g., busing consortia
representatives and local residents) can be lacking
\citep{buttazzoniSupportingActiveSchool2018}. In this sense, framing can
be used to position existing solutions as suitable to address particular
issues \citep{mahFramecriticalPolicyAnalysis2014}, which may prevent the
public from being aware of other policy approaches that challenge the
status quo. Contrariwise, framing can be instrumental to revealing more
diverse policy options than conventional wisdom would have
\citep{bosomworth2015climate}. Ultimately, it is important to understand
framing of policy issues because creating meaning through frames
contributes to the constant process of shaping of our collective
understanding of issues, uncovering what we collectively find
(un)acceptable and why \citep{beland2014developing}.

Given the role of education and encouragement in STP, it is important
for the issues to be effectively framed. In the past, parents have been
found to express understandings, language, and perceptions regarding the
influence of the built environment on school travel that markedly differ
from those of planners: while parents tend to view mixed land use as
conducive for driving, transport planners see mixed uses as a key for
encouraging more active travel \citep{buliungLivingJourneySchool2021}.
Similarly mismatched understandings have been detected with respect to
factors like the convenience of different modes to school
\citep{langUnderstandingModalChoice2011}. As a consequence, STP
stakeholders must pay special attention to parents' and communities'
understanding of the decline of AST as a problem, which may affect their
receptivity to proposed solutions
\citep{buttazzoniSupportingActiveSchool2018}.

Framing is important, but not all frames are equally effective. Frames
can appeal to economic reasons, to science and logic, or to moral
obligation. For example, \citet{severson2015moral} report that overall
support for climate change policy increases more when scientific frames,
secular moral frames, and economic equity frames are used, compared to
economic efficiency or religious moral frames. Frames can also be
positive or negative, and there is evidence that negative framing may be
more effective at motivating individuals, for example, to change
transport modes to reduce CO2 emissions
\citep{waygoodCO2ValenceFraming2018}. As well, the effectiveness of the
framing depends on the psychological distance to the relevant issues:
how socially, temporally, geographically, or hypothetically close an
individual perceives themselves to the issues can affect their support
for various policies \citep{maiella2020psychological}. For this reason,
stakeholders involved in STP must make conscious choices about how
issues surrounding AST are communicated if the goal is to increase
support for and buy-in of AST policies.

@@

We conducted a search on Web of Science for scholarly papers on the
topic of school travel using Web of Science's Core Collection. The
search was conducted in the Winter of 2021 and the parameters used were
\{active OR walk* or cycl* or bicycl*\} AND \{``school travel''\}.
Initially, we found 322 papers, which we reduced to 250 after selecting
only papers in the fields of transportation, planning, urban studies,
geography, and public health. These fields were chosen because they had
the greatest volume of documents and/or were of disciplinary interest to
the authors. This list with 250 documents was then manually curated by
the authors to ensure that all documents were relevant. This was done by
checking the title and reading the abstract of the papers. For example,
a paper with the title ``Impact of automated photo enforcement of
vehicle speed in school zones'' \citep{quistberg2019impact} was excluded
as being tangential to our research; the abstract of another paper
revealed that it was mainly concerned with survey methods: ``Common
methods for measuring mode share include Hands Up surveys and family
surveys, but these require teacher and parental involvement.''
\citep{sersli2019getting}. After this process, our academic corpus
comprised 227 journal articles that were readied for analysis.

@@

Topic modelling was used to analyze the documents retrieved. This is a
natural language processing technique used to analyze text to identify
the language and concepts being communicated. This method is practical
for researchers working with large amounts of text because it can assist
and complement the manual coding of topics that would normally take
place to analyze or summarize textual data
\citep{jacobiQuantitativeAnalysisLarge2016}. We primarily use the
following \texttt{R} packages: \texttt{tidytext}
\citep{silge2016tidytext}, \texttt{topicmodels} \citep{grun2011topic},
\texttt{word2vec} \citep{wijffels2021word2vec}, and \texttt{wordcloud}
\citep{fellows2018wordloud}.

@@

This paper is an example of open and reproducible research that uses
only open software. The source document is an R markdown document.
Following best practices in spatial data science
\citep{brunsdon2020opening}, the code and data needed to reproduce our
research or conduct a similar analysis for other regions are available
in a publicly available repository\footnote{\url{https://github.com/desjae/AST-Framing-Analysis-Ontario}}.

@@

We analyzed word and document frequency for each corpus of text. Table
\ref{tab:word-table} shows the most frequent terms found in the
municipal, transportation consortia, school board, and academic
documents. Policy documents and academic papers reference \emph{active},
\emph{travel}, \emph{walking}, \emph{biking} or \emph{cycling}, and
\emph{students} more than other terms. Each corpus also has
\emph{safety} and \emph{traffic} as common words, which suggests that
these are common concerns. The word \emph{physical} is present in each
corpus, but this could refers to \emph{physical activity} or the
\emph{physical environment}. Furthermore, documents from STP stakeholder
groups discuss \emph{resources}, \emph{information}, and \emph{services}
about school travel. Unlike the academic papers, policy+practice
documents frequently include the words \emph{route} or \emph{routes}.
This could indicate the role of STP stakeholder groups in identifying
safe routes to school to share with parents or families. In the section
below, the context in which these terms are used is explored further.

The academic corpus differs from the policy documents in that
\emph{parents} and \emph{distance} are the second and third most common
terms. In addition, \emph{time}, \emph{factors}, \emph{environment}, and
\emph{age} are also identified in the academic papers. The prevalence of
these terms is consistent with an academic focus on exploring the
variables that influence travel to school. These words are absent from
the list of common words in policy+practice documents. Table
\ref{tab:word-table} indicates that the academic corpus discusses a
broader range of determinants of AST than the policy documents. The
number of references for each term in the academic papers is also
substantially higher due to the inclusion of more documents.

@@

Examination of document frequency reveals terms that are not present in
all policy documents. This suggests that although documents pertain to
school travel, not all stakeholders across Ontario disseminate
information about AST. We manually searched the policy+practice corpus
and found that 48\% of documents mention AST and 16\% mention STP. In
contrast, inclement weather and its impacts on busing is a common topic
addressed in school board and transportation consortia documents.

@@

Bigrams refer to pairs of consecutive words. We combined all
municipality, school board, and transportation consortia documents into
the policy+practice corpus. Figure \ref{fig:policy-visual} shows all of
the bigrams that occur more than 10 times in this corpus. This figure
highlights the main ideas that are presented to the public in each of
the policy corpora. The directional arrows indicate the arrangement of
the words (e.g., active travel and not travel active) and the colour
gradient of the arrows corresponds to the most frequently mentioned
pairs (e.g., bigrams with darker arrows are found in the corpus more
often).
